---
title: "Iris classification"
author: "Naruhiko"
date: "2/25/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rpart)
library(rpart.plot)
library(ggplot2)
```

## Classificating Iris Species

This project aimed to deploy a machine learning algorithm to classify a new observation entry.

The algorithm received as input just *sepal length* and *petal length*.

The results seems to be pretty acurate, given the position of the new observation in the plot.

## Decision tree
A decision tree is created with both variables and a subset of the original iris dataset.  
The user can choose from 20% up to all the data available in the iris dataset to train the model.  
This is the default tree, with 70% of the dataset used to create it.

```{r, echo = FALSE, fig.align='center'}
        smp_size <- floor(0.7 * nrow(iris))
        
        ## set the seed to make your partition reproducible
        set.seed(123)
        train_ind <- sample(seq_len(nrow(iris)), size = smp_size)
        train <- iris[train_ind, ]
        
        model_tree <- rpart(Species ~ Sepal.Length + Petal.Length, data = train, method = 'class')
        
        tree <- prp(model_tree, 
                         sub = "", type = 2, extra = 'auto', fallen.leaves = TRUE, cex = 1,
                         left = FALSE, box.palette = 'auto', border.col = "#DBDBDC", varlen = 0,
                         round = 0.5, split.font = 1, split.cex = 0.9, branch.col = "grey90", 
                         branch.lty = 1, branch.type = 5, branch.tweak = 0.3, ge = " ≥ ",
                         space = 0.5, split.yshift = -1, split.space = 0, split.box.col = NULL)
```


## New entry

With the sliders, the user can create a new observation (diamond shaped) and see where it would be placed in the scatter plot.

The script will automatically classify this new entry using the new tree, giving the color
according to the its classification.
```{r, echo = FALSE, fig.align='center'}

x.extra <- 4.5
y.extra <- 6.5
pred <- predict(model_tree, newdata = data.frame(Sepal.Length = x.extra, 
                                                 Petal.Length = y.extra),
                                                 type = 'class')

paleta <- c('#f26666' , "#a0a1a3", "#5ec752") #vermelho, cinza e verde
        
new_point_color <- ifelse(pred == 'virginica', paleta[3],
                          ifelse(pred == 'versicolor', paleta[2], paleta[1]))
        
scatter <- ggplot(iris, aes(x = Sepal.Length, y = Petal.Length)) +
              geom_point(aes(color = Species)) +
              scale_color_manual(values = paleta) +
              geom_point(aes(x.extra, y.extra),
                         color = new_point_color,
                         size = 5,
                         alpha = 0.6,
                         shape = 23) +
              theme(panel.grid = element_blank(),
                    panel.background = element_rect(fill = "white"),
                    title = element_text(family = "Lato", colour = "#4D4D4D"),
                    axis.text = element_text(family = "Lato", colour = "#4D4D4D", size = 8),
                    axis.text.x = element_text(vjust = 2),
                    legend.text = element_text(family = "Lato", colour = "#4D4D4D"),
                    axis.ticks = element_line(colour = "#8F8F8F"),
                    panel.border = element_blank(),
                    axis.line = element_line(color = "#ebebeb", size = 0.25),
                    strip.background = element_rect(fill = '#ebebeb'),
                    strip.text = element_text(family = "Lato", colour = "#474747", size = 9),
                    plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
                  )
print(scatter)
```

## Slide with Plot

With this we can see the power of a decision tree as a machine learning algorithm or, at least,
how easy it is to classify different iris species according to its flower caracteristics.

Hope you have enjoyed!

